---
title: "Data Adjustment for `r params$country` --- Countdown Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
  officedown::rdocx_document:
    reference_docx: report-template.docx
  pdf_document:
    toc: false
    latex_engine: lualatex
params:
  cache: NULL
  country: NULL
  adminlevel_1: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = TRUE,
  fig.width=7, 
  fig.height=5, 
  dpi=1200
)

library(cd2030.rmncah)
library(dplyr)
library(purrr)
library(reactable)
library(officedown)
library(officer)

cache <- params$cache

data <- cache$countdown_data
k_factors <- cache$k_factors
```

## Health Facility Data Adjustments {-}

### Background

Completeness of reporting affects analysis, especially if it is low or varies between 
years. Extreme outliers can have a large impact, especially on subnational numbers.
Several steps are necessary to obtain a clean data set for “endline” analysis, 
including adjusting for incomplete reporting and correcting for extreme outliers. 
These graphs show the impact on the numbers.

```{r}
adjustments <- data %>%
  generate_adjustment_values(adjustment = 'custom', k_factors = !!k_factors)
```

### Health facility data adjustment: Numerators - reported live births

```{r}
plot(adjustments,
     indicator = 'instlivebirths',
     title = 'Comparison of number of live births before and after adjustments for completness and outliers') +
  cd_report_theme()
```

### Health facility data adjustment: Numerators - reported Penta 1 doses

```{r}
plot(adjustments,
     indicator = 'penta1',
     title = 'Comparison of number of penta1 vaccination before and after adjustments for completness and outliers') +
  cd_report_theme()
```

#### 📝 Notes 

::: {custom-style="Style1"}

---

Consider the effect of the adjustment on live births in health facilities, and mention the difference in the number, as well as the percent increase in 2023; highlight the year with the greatest impact if there is one; interpret if the impact of the adjustment on coverage rates is is large or small

Make the same description and interpretations for penta1 vaccinations 

---

:::
