---
title: "Health System Performance for `r params$country` --- Countdown Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
  officedown::rdocx_document:
    reference_docx: report-template.docx
  pdf_document:
    toc: false
    latex_engine: lualatex
params:
  cache: NULL
  country: NULL
  adminlevel_1: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = TRUE,
  fig.width=7, 
  fig.height=5, 
  dpi=1200
)

library(cd2030.rmncah)
library(dplyr)
library(purrr)
library(reactable)
library(officedown)
library(officer)

cache <- params$cache
data <- cache$adjusted_data
```

## Health system performance assessment: indicators

### Background

Subnational analyses of health system inputs and service outputs are critical: districts and regions are key units of the health systems and their service delivery. This includes assessment of system inputs (health workforce, infrastructure) and outputs (use, coverage).

### National Health System Indicators

```{r}
national_metrics <- calculate_health_system_metrics(data, 'national')
plot_national_health_metric(national_metrics, 'density') +
  cd_report_theme()
```

### Subnational Health System Indicators

```{r}
subnational_metrics <- calculate_health_system_metrics(data, 'adminlevel_1')

plot(subnational_metrics, indicator = 'ratio_fac_pop',
        national_score = national_metrics$ratio_fac_pop,
        target = 2) +
  cd_report_theme()
```

```{r}
plot(subnational_metrics, indicator = 'ratio_hos_pop',
        national_score = national_metrics$ratio_hos_pop,
        target = 2) +
  cd_report_theme()
```

```{r}
plot(subnational_metrics, indicator = 'ratio_hstaff_pop',
        national_score = national_metrics$ratio_hstaff_pop,
        target = 23) +
  cd_report_theme()
```

```{r}
plot(subnational_metrics, indicator = 'ratio_bed_pop',
        national_score = national_metrics$ratio_bed_pop,
        target = 25) +
  cd_report_theme()
```

#### ðŸ“ Notes 

::: {custom-style="Style1"}

---

What can be said about the data quality?

Describe comparison with health system data in WHO database and other countries? Does this suggest under- or overreporting?

Is the pattern by region/province as expected?

Any reasons for unusual data?

---

:::

## Health System Comparisons

```{r}
comparison <- cache$calculate_health_system_comparison()
plot(comparison, indicator = 'cov_instdeliveries_hstaff',
        denominator = cache$maternal_denominator) +
  cd_report_theme()
```

```{r}
plot(comparison, indicator = 'ratio_opd_u5_hstaff') +
  cd_report_theme()
```

```{r}
plot(comparison, indicator = 'ratio_opd_u5_hos') +
  cd_report_theme()
```

```{r}
plot(comparison, indicator = 'ratio_ipd_u5_bed') +
  cd_report_theme()
```


#### ðŸ“ Notes 

::: {custom-style="Style1"}

---

The interpretation should focus on whether the obtained picture has validity. This may be posited if it confirms some expectations such as more developed regions doing better

Describe poor and good performing regions/provinces.

---

:::
