---
title: "Curative Health Services for `r params$country` --- Countdown Analysis"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
    highlight: kate
  officedown::rdocx_document:
    reference_docx: report-template.docx
  pdf_document:
    toc: false
    latex_engine: lualatex
params:
  cache: NULL
  country: NULL
  adminlevel_1: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  include = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = TRUE,
  fig.width=7, 
  fig.height=5, 
  dpi=1200
)

library(cd2030.rmncah)
library(dplyr)
library(purrr)
library(reactable)
library(officedown)
library(officer)

cache <- params$cache
```

## Curative health services: admission and case fatality rates among children under-5

### Background

Data on inpatient admissions among under-fives are indicators of access to curative services. In-patient mortality (case fatality rates) is an indicator of quality of care. 


```{r, fig.cap='Figure 6a: OPD service use by children and all ages, national'}
service_utilization <- cache$compute_service_utilization('national')
service_utilization %>% 
  filter_service_utilization_map('opd') %>% 
  plot() +
  cd_report_theme()
```

```{r, fig.cap='Figure 6b : map with OPD service use by children, by region'}
admin1_service_utilization <- cache$compute_service_utilization('adminlevel_1')
min_year <- min(admin1_service_utilization$year)
max_year <- robust_max(admin1_service_utilization$year)

years <- if (is.null(cache$mapping_years)) c(min_year, max_year) else cache$mapping_years
admin1_filtered <- admin1_service_utilization %>% 
  cache$filter_service_utilization(indicator = 'opd', map_years = years)
plot(admin1_filtered) +
  cd_report_theme()
```

#### 📝 Notes 

::: {custom-style="Style1"}

---

What can be said about the data quality for OPD visits? Is there consistency of reported numbers between years? What is the % of OPD visits that are under-5 and are they within an expected range of 15-40%?? 

What is the number of OPD visits per child per year during 2019-2023, is it increasing?  Is it lower than 1 visit per year, which is considered indicative of low access? What % of OPD visits are for children under 5? 

What can be said about the OPD visits per child per year by region/province in 2023? How large is the difference between top and bottom regions? 

---

:::

## Curative health services: admission and case fatality rates among children under-5

### Background

Data on inpatient admissions among under-fives are indicators of access to curative services. In-patient mortality (case fatality rates) is an indicator of quality of care. 

```{r, fig.cap='Figure 6c: Admissions per 100 children and case fatality rates per 100 admissions under-5, national'}
service_utilization %>% 
  filter_service_utilization_map('ipd') %>% 
  plot() +
  cd_report_theme()
```

```{r, fig.cap='Figure 6d : Admission rates per 100 children under-5, by region'}
admin1_filtered <- admin1_service_utilization %>% 
  cache$filter_service_utilization(indicator = 'ipd', map_years = years)
plot(admin1_filtered) +
  cd_report_theme()
```

#### 📝 Notes 

::: {custom-style="Style1"}

---

What can be said about the data quality? Is there consistency of reported numbers of admissions / admission rates over time? Is the percent of admissions that are children under-5 plausible (within an expected range 15-40%) 

What is the number of admission per 100 children under 5 per year during 2019-2024?  Trend? Is it low or high?
What is the case fatality rate among admissions under-5, what is the trend? What does this say about the quality of care?

What can be said about admissions per 100 children under-5 per year by region/province in 2024?

---

:::
